---
title: "Factorial Analysis: a practical example"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

This blank document initialize the application of the factorial analysis previously displayed. You have the possibility to work on another dataset, the path and the file are already set.<br><br>

__Question__: Can you extract relevant correlations between variables ?<br>
__Question__: Can you extract the group of data with higher values of Y3 ?<br>
__Question__: Is it possible to characterize them as a function of X1, X2, ... ?


```{r, include=FALSE, echo=FALSE, message='hide'}
library(FactoMineR)
library(factoextra)
library(reticulate)
library(ggplot2)
library(ggpubr)
```
<br><br>
```{r}
current_path = "/media/marduq/One Touch/Personnel/BIG-MAP/AIWorkshop/"
X = read.csv(paste(current_path, "resources/forcefields-dataset.csv", sep=""), sep=",")
```

```{r, fig.height = 8, fig.width = 10}
res = PCA(X, scale.unit = TRUE, ncp = 5, graph = FALSE)
var.p = fviz_pca_var(res, axes=c(1, 2), col.var = "cos2", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE)
ggpar(var.p, title = "1st Principal Plan", font.x = "bold", font.y = "bold", font.title = "bold", subtitle = "",
      caption = "Source: factoextra/ggpubr", xlab = "Principal Component n°1", ylab = "Principal Component n°2",
      legend.title = "Cos2", legend.position = "top", ggtheme = theme_gray())
```